---
title: 'Case Study #2: Finacial Report (2015-2017)'
author: "Devin Masiak"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
data <- read_csv("casestudy.csv")
```

```{r functions, include=FALSE}
total_customers <- function(data, current_year) {
  if (is.na(current_year))
  {
    return(NA)
  }
  output_tibble <- data %>%
    filter(year==current_year) %>%
    summarize(customers = n())
  return(formatC(output_tibble$customers, format="f", digits=0, big.mark=","))
}

total_revenue <- function(data, current_year) {
  if (is.na(current_year))
  {
    return(NA)
  }
  output_tibble <- data %>%
    filter(year==current_year) %>%
    summarize(total_rev = sum(net_revenue))
  return(formatC(output_tibble$total_rev, format="f",digits=0, big.mark=","))
}

new_customer_total <- function(data, current_year, previous_year) {
  if (is.na(previous_year))
  {
    return(total_customers(data, current_year))
  }
  output_tibble <- data %>%
    filter(year==current_year | year==previous_year) %>%
    group_by(customer_email) %>%
    mutate(count = n()) %>%
    ungroup() %>%
    filter(year==current_year, count==1) %>%
    summarize(new_customers = n())
  return(formatC(output_tibble$new_customers, format="f", digits=0, big.mark=","))
}

new_customer_rev <- function(data, current_year, previous_year) {
  if (is.na(previous_year))
  {
    return(total_revenue(data, current_year))
  }
  output_tibble <- data %>%
    filter(year==current_year | year==previous_year) %>%
    group_by(customer_email) %>%
    mutate(count = n()) %>%
    ungroup() %>%
    filter(year==current_year, count==1) %>%
    summarize(new_rev = sum(net_revenue))
  return(formatC(output_tibble$new_rev, format="f", digits=0, big.mark=","))
}
# returns list of existing rev for both years
existing_rev <- function(data, current_year, previous_year) {
  if (is.na(previous_year))
  {
    return(c(NA,NA))
  }
  output_tibble <- data %>%
    filter(year==current_year | year==previous_year) %>%
    group_by(customer_email) %>%
    mutate(count = n()) %>%
    ungroup() %>%
    filter(count==2) %>%
    group_by(year) %>%
    summarize(existing_rev = sum(net_revenue))
  return(output_tibble$existing_rev)
}

lost_customer_total <- function(data, current_year, previous_year) {
    if (is.na(previous_year))
  {
    return(NA)
  }
  output_tibble <- data %>%
    filter(year==current_year | year==previous_year) %>%
    group_by(customer_email) %>%
    mutate(count = n()) %>%
    ungroup() %>%
    filter(year==previous_year, count==1) %>%
    summarize(lost_customers = n())
  return(formatC(output_tibble$lost_customers, format="f", digits=0, big.mark=","))
}

lost_customer_rev <- function(data, current_year, previous_year) {
  if (is.na(previous_year))
  {
    return(NA)
  }
  output_tibble <- data %>%
    filter(year==current_year | year==previous_year) %>%
    group_by(customer_email) %>%
    mutate(count = n()) %>%
    ungroup() %>%
    filter(year==previous_year, count==1) %>%
    summarize(lost_rev = sum(net_revenue))
  return(formatC(output_tibble$lost_rev, format="f", digits=0, big.mark=","))
}
```


```{r, include=FALSE}
current_year <- 2015
prev_year <- NA
yearly_revenue <- data %>% existing_rev(current_year,prev_year)
```
## 2015
Total Revenue = `r data %>% total_revenue(current_year)`

New Customer Revenue = `r data %>% new_customer_rev(current_year, prev_year)`

Existing Customer Growth = `r NA #formatC throws error on NA`

Revenue Lost from Attrition = `r NA #formatC throws error on NA`

Existing Customer Revenue Current Year = `r NA #formatC throws error on NA`

Existing Customer Revenue Prior Year = `r NA #formatC throws error on NA`

Total Customers Current Year = `r data %>% total_customers(current_year)`

Total Customers Previous Year = `r data %>% total_customers(prev_year)`

New Customers = `r data %>% new_customer_total(current_year, prev_year)`

Lost Customers = `r data %>% lost_customer_total(current_year, prev_year)`


```{r, include=FALSE}
current_year <- 2016
prev_year <- 2015
yearly_revenue <- data %>% existing_rev(current_year,prev_year)
```
## 2016
Total Revenue = `r data %>% total_revenue(current_year)`

New Customer Revenue = `r data %>% new_customer_rev(current_year, prev_year)`

Existing Customer Growth = `r formatC(yearly_revenue[2]-yearly_revenue[1], format="f", digits=0, big.mark=",")`

Revenue Lost from Attrition = `r formatC(as.numeric(gsub(',', '', data %>% lost_customer_rev(current_year, prev_year))) - as.numeric(gsub(',', '', data %>% new_customer_rev(current_year, prev_year))), format="f", digits=0, big.mark=",")`

Existing Customer Revenue Current Year = `r formatC(yearly_revenue[2], format="f", digits=0, big.mark=",")`

Existing Customer Revenue Prior Year = `r formatC(yearly_revenue[1], format="f", digits=0, big.mark=",")`

Total Customers Current Year = `r data %>% total_customers(current_year)`

Total Customers Previous Year = `r data %>% total_customers(prev_year)`

New Customers = `r data %>% new_customer_total(current_year, prev_year)`

Lost Customers = `r data %>% lost_customer_total(current_year, prev_year)`

```{r, include=FALSE}
current_year <- 2017
prev_year <- 2016
yearly_revenue <- data %>% existing_rev(current_year,prev_year)
```
## 2017
Total Revenue = `r data %>% total_revenue(current_year)`

New Customer Revenue = `r data %>% new_customer_rev(current_year, prev_year)`

Existing Customer Growth = `r formatC(yearly_revenue[2]-yearly_revenue[1], format="f", digits=0, big.mark=",")`

Revenue Lost from Attrition = `r formatC(as.numeric(gsub(',', '', data %>% lost_customer_rev(current_year, prev_year))) - as.numeric(gsub(',', '', data %>% new_customer_rev(current_year, prev_year))), format="f", digits=0, big.mark=",")`

Existing Customer Revenue Current Year = `r formatC(yearly_revenue[2], format="f", digits=0, big.mark=",")`

Existing Customer Revenue Prior Year = `r formatC(yearly_revenue[1], format="f", digits=0, big.mark=",")`

Total Customers Current Year = `r data %>% total_customers(current_year)`

Total Customers Previous Year = `r data %>% total_customers(prev_year)`

New Customers = `r data %>% new_customer_total(current_year, prev_year)`

Lost Customers = `r data %>% lost_customer_total(current_year, prev_year)`

## Dataset Plots
An interesting observation about this dataset is that the distribution of net revenue from customers is largely uniform. The only exceptions occur at the extreme ends of the range. I assume this is most likely a result of the data being randomly generated.
```{r, echo=FALSE, message=FALSE}
data %>%
  ggplot(aes(x=net_revenue, label=year)) +
  geom_histogram() +
  facet_wrap(~year) +
  labs(
      title = "Distribution of Customer Net Revenue",
      x = "Net Revenue",
      y = "Number of Customers"
    ) +
  theme_minimal()
```

It's also interesting to look at distribution of how long customers stay with the company. Only ~11% of customers remained all 3 years, while a whopping ~62% only signed up for a single year.

```{r, echo=FALSE, message=FALSE}
total <- nrow(data)
data %>%
  group_by(customer_email) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  group_by(count) %>%
  summarize(perc=n()/total*100) %>%
  ggplot(aes(x = as.factor(count), y = perc)) +
  geom_bar(stat = "identity") +
  labs(
      title = "Customer Retention",
      caption = "Data from 2015 to 2017",
      x = "Number of years",
      y = "Percent"
    ) +
  theme_minimal()
```